---
title: "APSTA-GE 2123 Assignment 4"
author: "Your Name"
output: 
  pdf_document: 
    number_sections: yes
---

# Oregon Medicaid Experiment

<!-- This is a chunk of R code -->

```{r}
J <- 50000 # number of households
dataset <- data.frame(household_ID = as.factor(unlist(lapply(1:J, FUN = function(j) {
  rep(j, each = sample(1:3, size = 1, prob = c(0.5, 0.3, 0.2)))
}))))
selection <- rbinom(nrow(dataset), size = 1, prob = 0.2)
dataset$lottery <- ave(selection, dataset$household_ID, FUN = any)
dataset$numhh <- as.factor(ave(dataset$lottery, dataset$household_ID, FUN = length))
```

## Actual Prior Predictive Distribution

The general functions for predicting income should be 
$$Income= \beta_{lottery}Lottery+\beta_{small}Small+\beta_{medium}Medium+\beta_{large}Large+\epsilon$$
```{r,message=FALSE}
rstan::expose_stan_functions(file.path('quantile_functions.stan'))
source(file.path('GLD_helpers.R'))
library(dplyr)
#distribution for household size of 1
beta_s_small<- GLD_solver_bounded(bounds=3000:100000,median=14700,IQR=3000)
#distribution for household size of 2
beta_s_medium<- GLD_solver_bounded(bounds=3000:100000,median=15000,IQR=3000)
#distribution for household size of 3 or above
beta_s_large<- GLD_solver_bounded(bounds=3000:100000,median=16000,IQR=3000)
#distribution for winning the lottery
beta_s_lottery<- GLD_solver_bounded(bounds=-1500:2000, median=-20, IQR=150)
#sigma for error
a_s_sigma<- GLD_solver(lower_quartile = 250 ,median=300, upper_quartile = 500, other_quantile = 0, alpha=0 )

#coefficient of household of size 1
beta_small<- GLD_rng(median=14700,IQR=3000,asymmetry = beta_s_small[1],steepness = beta_s_small[2])

#coefficient of household with size 2
beta_medium<- GLD_rng(median=15000, IQR=3000, asymmetry = beta_s_medium[1],steepness = beta_s_medium[2])

#coefficient of household with size 3 or above
beta_large<- GLD_rng(median=16000, IQR=3000, asymmetry = beta_s_large[1],steepness=beta_s_large[2])

#coefficient of winning the lottery
beta_lottery<- GLD_rng(median=-20, IQR=150, asymmetry = beta_s_lottery[1],steepness = beta_s_lottery[2])

#sigma for error to the estimation
sigma_<- GLD_rng(median=300, IQR=250, asymmetry = a_s_sigma[1],steepness = a_s_sigma[2])

#vector space for storing coefficient for different household size
gamma<- cbind(beta_small,beta_medium,beta_large)

dataset$income<- beta_lottery*dataset$lottery+gamma[dataset$numhh]+sigma_

#verify prediction on income
winning_lottery<- dataset %>% filter(lottery==1) %>% select(income)
summary(winning_lottery)

notwinning_lottery<- dataset %>% filter(lottery==0) %>% select(income)
summary(notwinning_lottery)
```


## Prior Predictive Distribution for a Journal

```{r}
#distribution for household size of 1
beta_s_small<- GLD_solver_bounded(bounds=3000:100000,median=14700,IQR=3000)
#distribution for household size of 2
beta_s_medium<- GLD_solver_bounded(bounds=3000:100000,median=15000,IQR=3000)
#distribution for household size of 3 or above
beta_s_large<- GLD_solver_bounded(bounds=3000:100000,median=16000,IQR=3000)
#refit distribution for winning the lottery with the median of 0
beta_s_lottery<- GLD_solver_bounded(bounds=-1480:2020, median=0, IQR=125)
#sigma for error
a_s_sigma<- GLD_solver(lower_quartile = 250 ,median=300, upper_quartile = 500, other_quantile = 0, alpha=0 )

#coefficient of household of size 1
beta_small<- GLD_rng(median=14700,IQR=3000,asymmetry = beta_s_small[1],steepness = beta_s_small[2])

#coefficient of household with size 2
beta_medium<- GLD_rng(median=15000, IQR=3000, asymmetry = beta_s_medium[1],steepness = beta_s_medium[2])

#coefficient of household with size 3 or above
beta_large<- GLD_rng(median=16000, IQR=3000, asymmetry = beta_s_large[1],steepness=beta_s_large[2])

#refit coefficient of winning the lottery with the median of 0
beta_lottery<- GLD_rng(median=0, IQR=125, asymmetry = beta_s_lottery[1],steepness = beta_s_lottery[2])

#sigma for error to the estimation
sigma_<- GLD_rng(median=300, IQR=250, asymmetry = a_s_sigma[1],steepness = a_s_sigma[2])

#vector space for storing coefficient for different household size
gamma<- cbind(beta_small,beta_medium,beta_large)

dataset$income<- beta_lottery*dataset$lottery+gamma[dataset$numhh]+sigma_

#verify prediction on income
winning_lottery<- dataset %>% filter(lottery==1) %>% select(income)
summary(winning_lottery)

notwinning_lottery<- dataset %>% filter(lottery==0) %>% select(income)
summary(notwinning_lottery)
```


# $2018$ American Community Survey

```{r, message = FALSE}
dataset <- readr::read_csv(dir(pattern = "csv$"))
dataset <- dataset[ , !startsWith(colnames(dataset), prefix = "PWG")]
dataset <- dataset[ , !startsWith(colnames(dataset), prefix = "F")]
dataset <- dataset[!is.na(dataset$WAGP) & dataset$WAGP > 0, ]
```

## Posterior Distribution
The following posterior distribution are performed with dataset on Nebraska. 
```{r, WAGP, cache = TRUE, results = "hide"}
library(rstanarm)
post_wgap<- stan_lm(log(WAGP)~AGEP+MAR+JWRIP+log(PINCP),data=dataset, prior=R2(location=0.65, what='mode'), adapt_delta=0.95)
```

```{r}
print(post_wgap,digits=4)
#To check plot and diagnostic use the following code
#launch_shinystan(post_wgap) 
```
Based on the information, we can conclude that Log of Personal earning variable has a positive coeffcient with relative high posterior distribution probability.


## Influential Observations

```{r}
library(rstanarm)
plot(loo(post_wgap),label_points=T,)
```
Based on the plot, observations 17295 seems to have an outsized influence on the posterior distributions. 

## Posterior Predictions

```{r}
# make histogram
Post_pred<- posterior_predict(post_wgap,draws=100,fun = exp)
pred_df<- as.data.frame(Post_pred)
pred_df$mean<- rowMeans(pred_df,na.rm=T)
hist(pred_df$mean,main="Average wages among wage-earners distribution",xlab="Wages",breaks=10)
```
Overall, there exists some uncertainty for people who's average wages are range from 47500 to 47600. As the shape of the distribution has a sudden drop instead of a concave bell-shape. 


## Topcoding

```{r}
topcoded_value <- max(dataset$WAGP)
# do the analysis
top_code_df<- dataset %>% filter(WAGP==430000) %>% select(AGEP,MAR,JWRIP,PINCP) %>% na.omit()
post_pred_top_code<- posterior_predict(post_wgap,newdata=top_code_df,draws=100,fun = exp)
exp_df<- as.data.frame(post_pred_top_code)
top_code_df$expectation_income<- colMeans(exp_df)
top_code_df$WAGP<- 430000
top_code_df
```
The posterior expectation for their actual income are recorded in the expectation_income column in the top_code dataframe.
